---
title: "MLR"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r load libraries, message=FALSE, warning=FALSE, echo=FALSE}

library(tidyverse)
library(prettydoc)
library(stringr)
library(stringi)
library(car)
require(glmnet)
library(fmsb)
library(broom)


```

# Predicting the Interest Rates using (Funded Amount,Term,Employment Title, Employment Length, Home Ownership, Annual Income, Purpose, Debt, Public Record)

```{r, echo=FALSE}
# Preprocessing
loan <- read.csv("../loan.csv")

#loan.new <- filter(loan, stri_detect_regex(loann$emp_title, "^C[EOMAIf]O$|^c[eomaif]o$|[[:alnum:]]*\\s*president|[[:alnum:]]*\\s*President|[[:alnum:]]*\\s*director|[[:alnum:]]*\\s*Director|[[:alnum:]]*\\s*Manager|[[:alnum:]]*\\s*manager|[[:alnum:]]*\\s*associate[^s]$|[[:alnum:]]*\\s*Associate[^s]$|[[:alnum:]]*\\s*analyst|[[:alnum:]]*\\s*Analyst|^vp$|^VP$"))

# Cleaning Up Employee Title
loan.new <- loan %>% 
  mutate(emp_title = case_when(
    stri_detect_regex(emp_title, "^C[EOMAIf]O$|
               ^c[eomaif]o$|
               [[:alnum:]]*\\s*president|
               [[:alnum:]]*\\s*President|
               [[:alnum:]]*\\s*director|
               [[:alnum:]]*\\s*Director|
               ^vp$|
               ^VP$") ~ "Top",
    stri_detect_regex(emp_title, "[[:alnum:]]*\\s*Manager|
               [[:alnum:]]*\\s*manager") ~ "Mid",
    stri_detect_regex(emp_title, "[[:alnum:]]*\\s*associate[^s]$|
               [[:alnum:]]*\\s*Associate[^s]$|
               [[:alnum:]]*\\s*analyst|
               [[:alnum:]]*\\s*Analyst") ~ "Low",
    TRUE ~ "Other"
  ))

# New Data Set
loan.new <- loan.new[,-c(1,2,3,5,8,9,10,15,16,17,18,19,20,22,23,24,26,27,28,29,30,31,33,34:74)]

## 75% of the sample size
smp <- floor(0.75 * nrow(loan.new))

## set the seed to make your partition reproducible
set.seed(123)
train_ind <- sample(seq_len(nrow(loan.new)), size = smp)

#Define Training and Test Set
train <- loan.new[train_ind, ]
test <- loan.new[-train_ind, ]

# Multiple Linear Regression Model
lm <- lm(data = train, int_rate~.)
summary(lm)

# Check for Multicollinearity
vif(lm);VIF(lm)

# Check for Normality
qqnorm(loan$int_rate);qqline(loan$int_rate)

# Removing Excess NAs
train <- train[complete.cases(train),]

# Regularization (Using LASSO)
m <- model.matrix(~., train)
y <- m[,4]
x <- m[,-4]

#Fitting the Model, getting the optimum lambda
lasso <- glmnet(x, y, alpha = 1)
cv.fit <- cv.glmnet(x,y,alpha = 1)
plot(cv.fit)
opt_lambda <- cv.fit$lambda.min
opt_lambda

#Extracting the best model
fin <- cv.fit$glmnet.fit
fin.coef <- coef(fin, s = opt_lambda)
y_predicted <- predict(fin, s = opt_lambda, newx = x)
sst <- sum((y - mean(y))^2)
sse <- sum((y_predicted - y)^2)
rsq <- 1 - sse / sst

# R-Squared
rsq

lasso.new <- glmnet(x,y,alpha = 1, lambda = opt_lambda)

# Influential Features
coef(lasso.new) %>%
  tidy() %>%
  filter(row != "(Intercept)") %>%
  ggplot(aes(value, reorder(row, value), color = value > 0)) +
  geom_point(show.legend = FALSE) +
  ggtitle("Influential variables") +
  xlab("Coefficient") +
  ylab(NULL)

```


**Multiple Regression in itself doesn't do great - 0.26**

**Penalized Multiple Regression with Lasso doesn't do great either - 0.26**

**Not Much Multicollinearity**

**Data not too normal**

**Data not linear Either**

**Next Steps - Transform Data, use nonparametric regression, random forests and other ensemble methods**


```{r, eval=FALSE, echo=FALSE}

# Multiple Regression Model

lm <- lm(data = train, int_rate~.)
summary(lm)

# Check for Multicollinearity
vif(lm)
VIF(lm)
cor(loan$funded_amnt, loan$installment)
cor(train)

train <- train[,-7]

# Redo Multiple regression after removing "installment"

lm <- lm(data = train, int_rate~.)


# Residuals vs Fitted

ggplot(lm,aes(.fitted, .resid)) + geom_point() + stat_smooth(method = "auto") +
  geom_hline(yintercept=0, col="red", linetype="dashed") + 
  xlab("Fitted values")+ylab("Residuals") + ggtitle("Residual vs Fitted Plot") + theme_bw()

# Residuals vs Fitted for each variable

train$resid <- residuals(lm)
library(reshape2)
plotDF <- melt(train[, c("emp_length", "grade", "purpose","annual_inc","installment",
                         "dti","application_type","addr_state","empl_title", "resid")], id="resid")
library(ggplot2)
ggplot(plotDF, aes(x=value, y=resid)) +  geom_hline(yintercept=0, col="red", linetype="dashed") +
  geom_point() + stat_smooth(method = "auto") + facet_wrap(~ variable, scales = "free")


# Check for Normality

qqnorm(train$funded_amnt); qqline(train$funded_amnt)
ks.test(train$funded_amnt, pnorm)

```


```{r, echo=FALSE, eval=FALSE}

emp.type <- split(loan.new, loan.new$empl_title)

# Top Employees

p <-  ggplot(emp.type$Top,aes(x=as.factor(grade),  
                    y=as.factor(purpose), 
                    color=as.factor(grade))) + 
  ggtitle("Top employees get which loan grade for what purpose") +
  xlab("Grade") + ylab("Purpose")

p + geom_jitter(alpha=0.3)

# Mid Employees

q <-  ggplot(emp.type$Mid,aes(x=as.factor(grade),  
                    y=as.factor(purpose), 
                    color=as.factor(grade))) + 
  ggtitle("Top employees get which loan grade for what purpose") +
  xlab("Grade") + ylab("Purpose")

q + geom_jitter(alpha=0.3)

# Low Employees

r <-  ggplot(emp.type$Low,aes(x=as.factor(grade),  
                    y=as.factor(purpose), 
                    color=as.factor(grade))) + 
  ggtitle("Top employees get which loan grade for what purpose") +
  xlab("Grade") + ylab("Purpose")

r + geom_jitter(alpha=0.3)

```


**In which state which type of employee gets a loan for what purpose?**

```{r, fig.width=15, fig.height=9, echo=FALSE, eval=FALSE}

p <-  ggplot(loan.new,aes(x=as.factor(addr_state),  
                    y=as.factor(purpose), 
                    color=as.factor(home_ownership))) + 
  ggtitle("Statewise: Which type of employee gets loans for what purposes?") +
  xlab("state") + ylab("Purpose")

p + geom_jitter(alpha=0.3)



```

*Interesting to see that CA folks get loans for cars, major purchases, moving and medical apart from the common debt consolidation and credit card)*

*Mortgagers take loans a lot more than home owners for home improvement*


# We can slice the data by home_ownership/Employee Type/Application type
# Dig deeper into one state
# Dig Deeper into one grade


